---
title: "Ice Cream Analysis"
author: "Harris Temuri"
date: "4/7/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Time Series Analysis of Ice Cream Search Popularity on Google.com

What is a time series? A time series is a series of data points taken with respect to time. The time series data that I chose to analyze is a series of Ice Cream search popularity on the Google search engine.

What is search popularity? According to Google, "Numbers represent search interest relative to the highest point on the chart for the given region and time. A value of 100 is the peak popularity for the term. A value of 50 means that the term is half as popular. A score of 0 means there was not enough data for this term."

### Goal of the Project

Using the dataset provided by Google, I will create a model to predict the search popularity of Ice Cream. I will use data beginning from January 1, 2004 to December 1, 2020 to generate the model. Afterwards, I will compare the predicted values for the months of January, 2021 - April 1, 2021 to the actual values from Google.

### Initializing the project

Install and Load the following libraries:

```{r}
library(readr)
library(MASS)
library(ggplot2)
library(dplyr)
library(TSstudio)
library(forecast)
library(tseries)
```

Next, I will create a time series of the "training" data and another for the entire data set

```{r}
fullData <- read_csv("iceCreamData.csv",col_types = cols(Month = col_date(format = "%Y-%m")))

iceCream <- ts(fullData[1:204, 2], freq=12, start=c(2004, 1))
iceCream.full <- ts(fullData[, 2], freq=12, start=c(2004, 1))
```

### Ice Cream Search Popularity Time Series Plot

```{r}
# Plot Time Series
ts_plot(iceCream,
        title = "Ice Cream Search Popularity",
        Xtitle = "Time",
        Ytitle = "Interest Over Time"
)
```

We can also take a look at the acf and pacf graphs of the raw data:

```{r}
# Plotting acf and pacf before data manipulation
par(mfrow=c(2,1))
acf(iceCream, lag.max = length(iceCream))
pacf(iceCream, lag.max = length(iceCream))
```
### Decomposing Series

Looking at these plots, we can clearly tell that there is a strong seasonality. In order to fit an ARIMA model to the time series, we need to remove the seasonality. To remove the seasonality and decompose the series, I will use the stl function which comes with the built in the stats package in R. I will then create a series of just the trend and residuals.

```{r}
# Decomposing series
iceCream.stl <- stl(iceCream[, 1], "periodic")
iceCream.full.stl <- stl(iceCream.full[, 1], "periodic")
plot(iceCream.stl)
```

The data is now decomposed into a trend, seasonality, and residuals. I will create a new series with just the trend and random error components of the series and show the new plot.

```{r}
# Removing Seasonality
iceCream.series <- iceCream.stl$time.series[,2] + iceCream.stl$time.series[, 3]

# Plot Time Series
ts_plot(iceCream.series,
        title = "Ice Cream Search Popularity Deseasonalized",
        Xtitle = "Time",
        Ytitle = "Interest Over Time"
)

# Plotting acf and pacf after removing seasonality
par(mfrow=c(2,1))
acf(iceCream.series, lag.max = length(iceCream.series))
pacf(iceCream.series, lag.max = length(iceCream.series))
```

Now that the series is deseasonalized, we need to decide what orders to use for the ARIMA model. There are a few different ways of determining this, but I will choose my model based on the AIC value. Since the data was not differenced, my "d" value is going to remain 0.

### AIC to determine best ARIMA order

```{r}
# Determining best ARIMA Model for the Time Series by comparing AIC values
aicMatrix <- matrix(nrow = 4, ncol = 4)
for (r in 1:4) {
  for (l in 1:4) {
    #print(paste("(",r,",",l,")"))
    aicMatrix[r,l] <- AIC(arima(iceCream.series, order=c(r,0,l)))
    #print(aicMatrix[r,l]) 
  }
}

aicTable <- as.table(aicMatrix)
colnames(aicTable) <- c(1,2,3,4)
rownames(aicTable) <- c(1,2,3,4)
aicTable

```

Looking at the table of AIC values at different orders, we get the lowest AIC from ARIMA(2,0,4)

```{r}
# We will use ARIMA(2,0,4) because that gave us the lowest AIC
iceCream.series.arma1 <- arima(iceCream.series, order=c(2,0,4))
iceCream.series.arma1

# Plotting acf and pacf of residuals
par(mfrow=c(2,1))
acf(iceCream.series.arma1$residuals, lag.max = length(iceCream.series.arma1$residuals))
pacf(iceCream.series.arma1$residuals, lag.max = length(iceCream.series.arma1$residuals))

# Testing with first 10 lags
checkresiduals(iceCream.series.arma1, lag = 12)
```

### Try Again with different decomposition

The residuals show up as dependent, so we cannot trust the resulting model predictions. I will use the mstl function instead to decompose and deseasonalize the data.

```{r}
# Decomposing series
iceCream.new.stl <- iceCream %>% mstl()

# Removing Seasonality
iceCream.new.series <- iceCream.new.stl[,2] + iceCream.new.stl[, 4]

# We will use ARIMA(4,0,4) because that gave us the lowest AIC
iceCream.new.series.arma1 <- arima(iceCream.new.series, order=c(4,0,4))

# Analyzing Residuals
checkresiduals(iceCream.new.series.arma1)
```

Our residuals are now normal, so we can believe in the data. Finally, we have make a prediction data series and compare it to our sample data.

### Prediction data and Final Results

```{r}
# Prediction of the series
iceCream.new.pred <- predict(iceCream.new.series.arma1, n.ahead = 4)
# Actual Data
iceCream.future <- ts(fullData[205:208,2], frequency = 12, start = c(2021,1))
# Seasonality of actual data
iceCream.future.seasonality <- iceCream.full.stl$time.series[205:208,1]
# Final prediction data
newFinalPred <- iceCream.new.pred$pred + iceCream.new.pred$se
# Deseasonalized actual data
ActualDeseasonalized <- iceCream.future - iceCream.future.seasonality

newFinalPred
ActualDeseasonalized

```
All of our predicted values of Ice Cream search popularity were within error of the actual values.